### 2023-04-09 main/chapter_experiment.tex
add beam size

\section{The Large Hadron Collider at CERN}
\label{sec:atlas_lhc}
% The LHC
% largest piece
The Large Hadron Collider (LHC) is the largest and most powerful piece of
CERN's extensive particle accelerator complex.
It lives in a $26.7\,\mathrm{km}$ (circumference) tunnel, which is buried
around $100\,\mathrm{m}$ under France and Switzerland at CERN, Geneva.
It and comprises two rings that contain and accelerate counter-rotating bunches
of particles.
These accelerated bunches are brought into near-head-on collisions at
the four large detector experiments:
\atlas~\cite{atlas2008experiment},
\lhcb~\cite{lhcb2008experiment},
\cms~\cite{cms2008experiment}, and
\alice~\cite{alice2008experiment}.
Although proton-proton collisions are the LHC's primary subject, it is also
operated for heavy ion collisions including lead ($^{208}\mathrm{Pb}^{82+}$)
and oxygen ($^{16}\mathrm{O}^{4+}$)~\cite{
Brewer:2021kiv,
lhc2008machine,
lhcDesignreport2004
}.
Heavy ion collisions are primarily studied by the \alice\ experiments, but are
also analysed by the general-purpose detectors \atlas\ and \cms.
Since proton collision data are the topic of this thesis, this section focuses
on proton operations.

@@

% higher orders
Sextupole, octupole, and decapole magnets are used to implement detailed
corrections with each of the dipoles magnets.
Chromatic aberrations in optics are errors that arise from the colour-dependent
behaviour of light.
The analogous momentum-dependent behaviour in particle beams is known as
`chromaticity'.
Sextupole magnets are used in the LHC to correct for chromaticity at each
quadrupole~\cite{Guiducci:1992kb, lhc2008machine}.

% TODO size of beam
% TODO luminosity and its uncertainty
% TODO pile-up
\TODO{Distribution in the mean number of primary vertices per crossing, where the
variability in that mean arises from differing beam conditions over run-time.}
Figure~\ref{fig:experiment_run_2_mu}

>>>>

\section{The Large Hadron Collider}
\label{sec:atlas_lhc}
% The LHC
% largest piece
The Large Hadron Collider (LHC) is the largest and most powerful piece of
CERN's extensive particle accelerator complex.
It lives in a $26.7\,\mathrm{km}$ (circumference) tunnel, which is buried
around $100\,\mathrm{m}$ under France and Switzerland at CERN, Geneva.
It and comprises two rings that contain and accelerate counter-rotating bunches
of particles.
These accelerated bunches are brought into near-head-on collisions at
the four large detector experiments:
\atlas~\cite{atlas2008experiment},
\lhcb~\cite{lhcb2008experiment},
\cms~\cite{cms2008experiment}, and
\alice~\cite{alice2008experiment}.
Although proton-proton collisions are the LHC's primary subject, it is also
operated for heavy ion collisions including lead ($^{208}\mathrm{Pb}^{82+}$)
and oxygen ($^{16}\mathrm{O}^{4+}$)~\cite{
Brewer:2021kiv,
lhc2008machine,
lhcDesignreport2004
}.
Heavy ion collisions are primarily studied by the \alice\ experiments, but are
also analysed by the general-purpose detectors \atlas\ and \cms.
Since proton collision data are the topic of this thesis, this section focuses
on proton operations.

@@

% higher orders
Sextupole, octupole, and decapole magnets are used to implement detailed
corrections with each of the dipoles magnets.
Chromatic aberrations in optics are errors that arise from the colour-dependent
behaviour of light.
The analogous momentum-dependent behaviour in particle beams is known as
`chromaticity'.
Sextupole magnets are also used in the LHC to correct for chromaticity at each
quadrupole~\cite{Guiducci:1992kb, lhc2008machine}.

% size of beam
By its peak luminosity design specifications, each bunch of an LHC beam
should contain $1.15\times10^{11}$ protons.
The ring can contain up to $2808$ bunches.
After acceleration to collision energy, each bunch has a root-mean-square
(RMS, standard-deviation) length of $7.55\,\mathrm{cm}$.
The beam has a transverse RMS width of $16.7\,\mathrm{\mu m}$ at
interaction points 1 and 5 (\atlas\ and \cms),
and of $70.9\,\mathrm{\mu m}$ at
interaction points 2 and 8 (\alice\ and \lhcb)%
~\cite{lhcDesignreport2004}.

% TODO luminosity and its uncertainty
% TODO pile-up
\TODO{Distribution in the mean number of primary vertices per crossing, where the
variability in that mean arises from differing beam conditions over run-time.}
Figure~\ref{fig:experiment_run_2_mu}
